---
title: "Tidycensus Markdown"
author: "Jiatong su"
date: september 5, 2024
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(pander)
library(rmarkdown)
```

# Setup 

Once the packages are installed, you must load them using hte `library` command so that they are active in your environment.

```{r setup_packages2, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidycensus)
library(sf)
```

## Census API Key


```{r load_key, warning = FALSE, eval = FALSE}
census_api_key("3f9a49c7b9734670b02759f6daeeb5f8c605be14", overwrite=TRUE)

#install = TRUE
```

## Load census data dictionaries


```{r load_variables, cache = TRUE}

acs_variable_list.2020 <- load_variables(2020, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

acs_variable_list.2016 <- load_variables(2016, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)
```


# Downloading Data from Tidycensus

## Create a vector of census variables

```{r acs_vars}
acs_vars <- c("B01001_001E", # ACS total Pop estimate
              "B25002_001E", # Estimate of total housing units
              "B25002_003E", # Number of vacant housing units
              "B19013_001E", # Median HH Income ($)
              "B02001_002E", # People describing themselves as "white alone"
              "B06009_006E") # Total graduate or professional degree
```

## Call the Census API to get tract level data for 2020 for all of Philadelphia


```{r get_acs_2020, cache = TRUE, message = FALSE, warning = FALSE}
acsTractsPHL.2020 <- get_acs(geography = "tract",
                             year = 2020, 
                             variables = acs_vars, 
                             geometry = FALSE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") 
```
# Wrangling Data with dplyr
## Mutating, selecting and renaming variables

```{r}
acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  dplyr::select (GEOID, NAME, all_of(acs_vars))

acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  rename (total_pop.2020 = B01001_001E,
          total_HU.2020 = B25002_001E,
          total_vacant.2020 = B25002_003E,
          med_HH_Income.2020 = B19013_001E,
          total_White.2020 = B02001_002E,
          total_GradDeg.2020 = B06009_006E)

acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  mutate(vacancyPct.2020 = total_vacant.2020/total_HU.2020,
         pctWhite.2020   = total_White.2020/total_pop.2020)
```

```{r}
acsTractsPHL.2016 <- get_acs(geography = "tract",
                             year = 2016, 
                             variables = acs_vars,
                             geometry = FALSE,
                             state = "PA", 
                             county = "Philadelphia",
                             output = "wide") %>%
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>% 
  rename (total_pop.2016 = B01001_001E,
          total_HU.2016 = B25002_001E,
          total_vacant.2016 = B25002_003E,
          med_HH_Income.2016 = B19013_001E,
          total_White.2016 = B02001_002E,
          total_GradDeg.2016 = B06009_006E) %>%
  mutate(vacancyPct.2016 = total_vacant.2016/total_HU.2016,
         pctWhite.2016 = total_White.2016/total_pop.2016)

```

# joining data
```{r}
allACS <- left_join(acsTractsPHL.2016, acsTractsPHL.2020,
                    by= c("GEOID"))
```


# Comparing geographies

```{r myTracts, cache = TRUE}
myTracts <- c("42101023500", 
              "42101023600", 
              "42101023700", 
              "42101025300", 
              "42101025400",
              "42101025500", 
              "42101025600", 
              "42101038800")

allACS <- allACS %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA"))
```



# Graphic comparisons Using ggplot (2a)
a ggplot plot city-wide - number of vacant housing units per tract in 2020 as a function of vacant housing units in 2016,  (change in raw housing units, so 2020-2016, using "B25002_003E", # Number of vacant housing units", which is part of the `acs_vars` vector). Use the ggplot "color" parameter in the geom_point aesthetics to visualize which are the Mt. Airy tracts.  Annotate this plot with a single sentence describing what it shows (not as a plot title, but within the markdown).

```{r ggplot_point2, warning = FALSE, cache = TRUE, message = FALSE}
## assign 2(A)
ggplot(allACS)+
  geom_point(aes(x = total_vacant.2016, 
                 y = total_vacant.2020,
                 color = mtAiry))+
  geom_abline(intercept = 0, slope = 1)+
  labs(
    title = "Vacant housing units in 2020 as a function of vacant housing units in 2016",
    caption = "Data: US Census Bureau, ACS 5-year estimates",
    x="total_vacant 2016 ", 
    y="total_vacant 2020 ")
  
```

# Spatial Data and Tidycensus (2b)
A ggplot map symbolizing the number of vacant housing units in 2016 as a percentage of total housing units (e.g.  "total_HU.2016") on a per-tract basis. Outline the Mt. Airy neighborhood on the map. Annotate this map with a single sentence describing what it shows or implies about Mt Airy and Philadelphia.

```{r spatial_tidycensus, message=FALSE, warning=FALSE, cache=TRUE, include=FALSE}
acsTractsPHL.2016.sf <- get_acs(geography = "tract",
                             year = 2016, 
                             variables = acs_vars, 
                             geometry = TRUE, 
                             state = "PA", 
                             county = "Philadelphia", 
                             output = "wide") %>% 
  dplyr::select (GEOID, NAME, all_of(acs_vars)) %>%
  rename (total_pop.2016 = B01001_001E,
          total_HU.2016 = B25002_001E,
          total_vacant.2016 = B25002_003E,
          med_HH_Income.2016 = B19013_001E,
          total_White.2016 = B02001_002E,
          total_GradDeg.2016 = B06009_006E) %>%
  mutate(vacancyPct.2016 = (total_vacant.2016/total_HU.2016)*100,
         pctWhite.2016 = total_White.2016/total_pop.2016) %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA")) %>%
  st_as_sf(crs = 4326) # Turn shp into sf object and project as WGS84
```


```{r ggplot_geom_sf, warning = FALSE, cache = TRUE}
ggplot()+
  geom_sf(data = acsTractsPHL.2016.sf, aes(fill = vacancyPct.2016),
          color = "transparent")+
  geom_sf(data = acsTractsPHL.2016.sf %>%
            filter(mtAiry == "MT AIRY") %>%
            st_union(),
          color = "white",
          fill = "transparent")+
  labs(
    title = "Percentage of those identifying as 'vacant housing' by tract",
    subtitle = "",
    caption = "Data: US Census Bureau, ACS 5-year estimates")
```


# Assignment 2c
Create a code chunk that renders a "kable" table showing a summary of **mean** number of vacant housing units per tract for Mt. Airy vs. the rest Philadelphia as a whole in 2020.
```{r}
#install.packages("kableExtra")
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
```

```{r}
acsTractsPHL.2020 <- acsTractsPHL.2020 %>%
  mutate(mtAiry = ifelse(GEOID %in% myTracts, "MT AIRY", "REST OF PHILADELPHIA"))
```

```{r}
summaryTable_2c <- acsTractsPHL.2020 %>%
  group_by(mtAiry) %>%
  summarize(mean_vacant = mean(total_vacant.2020, na.rm = TRUE))
```

```{r}
dt <- summaryTable_2c 
dt %>%
  kable("html") %>%
  kable_styling()
```
