---
title: "Lab 2: Why Start With Indicators?"
author: "Jiatong su"
date: '2024-09-12'
output: html_document
code_folding: show
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rm(list = ls())

```

# Load packages and functions

We begin by loading packages, loading our API key etc.,

```{r setup_packages, warning = FALSE, message = FALSE}
# Load Libraries

library(tidyverse)
library(tidycensus)
library(sf)
library(kableExtra)

options(scipen=999)
options(tigris_class = "sf")

source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

palette5 <- c("#f0f9e8","#bae4bc","#7bccc4","#43a2ca","#0868ac")
```

```{r load_key, warning = FALSE, eval = FALSE}
census_api_key("d9ebfd04caa0138647fbacd94c657cdecbf705e9", overwrite = TRUE)
```

### Use `get_acs()` to get 2016 ACS data

```{r}
tracts12 <-  
  get_acs(geography = "tract",
          variables = c("B25026_001E","B02001_002E",
                        "B15001_050E","B15001_009E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E"), 
          year=2012, state=11,
          county=001, geometry=TRUE) %>% 
  st_transform('ESRI:102728')
```

```{r results='hide'}
tracts12_long <-  
  get_acs(geography = "tract",
          variables = c("B25026_001E","B02001_002E",
                        "B01001_011E","B01001_012E","B01001_013E","B01001_014E","B01001_015E","B01001_016E","B01001_017E","B01001_018E","B01001_019E","B01001_020E","B01001_021E","B01001_022E","B01001_023E","B01001_024E","B01001_025E",
                        "B01001_035E","B01001_036E","B01001_037E","B01001_038E","B01001_039E","B01001_040E","B01001_041E","B01001_042E","B01001_043E","B01001_044E","B01001_045E","B01001_046E","B01001_047E","B01001_048E","B01001_049E",
                        "B15001_017E","B15001_025E","B15001_033E","B15001_041E",
                        "B15001_058E","B15001_066E","B15001_074E","B15001_082E",
                        "B19013_001E", "B25058_001E",
                        "B06012_002E","B25026_001"), 
          year=2012, state=11,
          county=001, geometry=TRUE) %>% 
  st_transform('ESRI:102728')
```

#### Wide data vs long data (and `spread`vs `gather`)

Let's examine each variable and the elements of an sf object

```{r}
glimpse(tracts12_long)
```

We create a new data frame consisting only of population

```{r}
totalPop12 <-
  tracts12_long %>%
  filter(variable == "B25026_001")
```

![wide vs long
data](https://github.com/mafichman/musa_5080_2024/blob/main/Week_2/images/tidyr-spread-gather.gif?raw=true)

Image: Garrick Buie

*Why do you think this data type might be useful? Can you name a couple
of examples?*

### Using ggplot to map census data with {sf} and quantile symbologies

*Exercise: Let's make a few plots by filtering `B25026_001` within a
ggplot statement and visualizing it. Each plot adds more and more nuance
and information. Run the code, and examine each to see what we've added
each time.*

We include a couple of custom functions from our textbook:

*What does `q5` do?*

*What does `qBr` do?*

(Note: the `qBr()` function used here is an update to the one used
version used in the text book so the results may appear different.)

Plot 1:

```{r}
ggplot() +
  geom_sf(data = tracts12_long %>% 
            filter(variable == "B25026_001"), 
          #Estimate!!Total population in occupied housing units:
          aes(fill = estimate)) +
  theme(
    plot.title = element_text(size=22)
    )
```

Plot 2:

```{r}
  ggplot() +
  geom_sf(data = tracts12_long %>% 
            filter(variable == "B25026_001"),
          aes(fill = q5(estimate))) +
  theme(plot.title = element_text(size=22)) 
```

Plot 3:

```{r}
ggplot() +
  geom_sf(data = tracts12_long %>% 
            filter(variable == "B25026_001"),
          aes(fill = q5(estimate))) +
  scale_fill_manual(values = palette5,
                    labels = qBr(tracts12 %>% 
            filter(variable == "B25026_001"), "estimate"),
                    name = "Total\nPopluation\n(Quintile Breaks)") +
  theme(plot.title = element_text(size=22))
```

Plot 4:

```{r}
ggplot() +
  geom_sf(data = tracts12_long %>% 
            filter(variable == "B25026_001"), 
          aes(fill = q5(estimate))) +
  scale_fill_manual(values = palette5,
                    labels = qBr(tracts12_long %>% 
            filter(variable == "B25026_001"), "estimate"),
                    name = "Population\n(Quintile Breaks)") +
  labs(title = "Total Population", subtitle = "Washington DC; 2012") +
  mapTheme() + 
  theme(plot.title = element_text(size=22))
```

As we can see the map, NW has the densest population, and SW has the
fewest population.

### Transforming Long Data to Wide Data using spread

Here we will modify the 2016 ACS by using `spread()` to widen the data -
each census variable is now a column. This looks like the data we had in
Lab 1.

We will use `rename()` to give sensible names to the columns and we will
use `mutate()` to make new features out of the existing columns. These
are all `dplyr` "verbs" we saw in Lab 1, except here we are doing it on
{sf} spatial data frames.

```{r}
tracts12_wide <- 
  tracts12_long %>%
  dplyr::select( -NAME, -moe) %>%
  spread(key = variable, value = estimate) %>%
  rename(TotalPop = B25026_001, 
         Whites = B02001_002,
         MPop25_29 = B01001_011,
         MPop30_34 = B01001_012,
         MPop35_39 = B01001_013,
         MPop40_44 = B01001_014,
         MPop45_49 = B01001_015,
         MPop50_54 = B01001_016,
         MPop55_59 = B01001_017,
         MPop60_61 = B01001_018,
         MPop62_64 = B01001_019,
         MPop65_66 = B01001_020,
         MPop67_69 = B01001_021,
         MPop70_74 = B01001_022,
         MPop75_79 = B01001_023,
         MPop80_84 = B01001_024,
         MPop85_plus = B01001_025,
         FPop25_29 = B01001_035,
         FPop30_34 = B01001_036,
         FPop35_39 = B01001_037,
         FPop40_44 = B01001_038,
         FPop45_49 = B01001_039,
         FPop50_54 = B01001_040,
         FPop55_59 = B01001_041,
         FPop60_61 = B01001_042,
         FPop62_64 = B01001_043,
         FPop65_66 = B01001_044,
         FPop67_69 = B01001_045,
         FPop70_74 = B01001_046,
         FPop75_79 = B01001_047,
         FPop80_84 = B01001_048,
         FPop85_plus = B01001_049,
         FBach25_34 = B15001_058,
         FBach35_44 = B15001_066,
         FBach45_64 = B15001_074,
         FBach65plus = B15001_082,
         MBach25_34 = B15001_017,
         MBach35_44 = B15001_025,
         MBach45_64 = B15001_033,
         MBach65plus = B15001_041,
         MedHHInc = B19013_001, 
         MedRent = B25058_001,
         TotalPoverty = B06012_002)
```

Let's create new rate variables using mutate..

```{r}
tracts12 <- 
  tracts12_wide %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         TotalMPop25plus = MPop25_29 + MPop30_34 + MPop35_39 + MPop40_44 + MPop45_49 + MPop50_54 + MPop55_59 + MPop60_61 + MPop62_64 + MPop65_66 + MPop67_69 + MPop70_74 + MPop75_79 + MPop80_84 + MPop85_plus,
         TotalFPop25plus = FPop25_29 + FPop30_34 + FPop35_39 + FPop40_44 + FPop45_49 + FPop50_54 + FPop55_59 + FPop60_61 + FPop62_64 + FPop65_66 + FPop67_69 + FPop70_74 + FPop75_79 + FPop80_84 + FPop85_plus,
         TotalPop25plus = TotalMPop25plus + TotalFPop25plus,
         FemaleBachelors = FBach25_34 + FBach35_44 + FBach45_64 + FBach65plus,
         MaleBachelors = MBach25_34 + MBach35_44 + MBach45_64 + MBach65plus,
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop25plus), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2012") %>%
  dplyr::select(GEOID,TotalPop,MedHHInc,MedRent,pctWhite,pctBachelors,pctPoverty,year)
```

Tracts 2012 is now complete. Let's grab 2024 tracts in "wide" format and
do the same renaming.

### 2023 Census Data

```{r results='hide'}
tracts22 <- 
  get_acs(geography = "tract", 
          variables = c("B25026_001E","B02001_002E",
  "B01001_011E","B01001_012E","B01001_013E","B01001_014E","B01001_015E","B01001_016E","B01001_017E","B01001_018E","B01001_019E","B01001_020E","B01001_021E","B01001_022E","B01001_023E","B01001_024E","B01001_025E",
  "B01001_035E","B01001_036E","B01001_037E","B01001_038E","B01001_039E","B01001_040E","B01001_041E","B01001_042E","B01001_043E","B01001_044E","B01001_045E","B01001_046E","B01001_047E","B01001_048E","B01001_049E",
  "B15001_017E","B15001_025E","B15001_033E","B15001_041E",
  "B15001_058E","B15001_066E","B15001_074E","B15001_082E",
  "B19013_001E", "B25058_001E",
  "B06012_002E"), 
          year=2022, 
          state=11, 
          county=001, 
          geometry=TRUE, 
          output="wide") %>%
  st_transform('ESRI:102728') %>%
  rename(TotalPop = B25026_001E, 
         Whites = B02001_002E,
         MPop25_29 = B01001_011E,
         MPop30_34 = B01001_012E,
         MPop35_39 = B01001_013E,
         MPop40_44 = B01001_014E,
         MPop45_49 = B01001_015E,
         MPop50_54 = B01001_016E,
         MPop55_59 = B01001_017E,
         MPop60_61 = B01001_018E,
         MPop62_64 = B01001_019E,
         MPop65_66 = B01001_020E,
         MPop67_69 = B01001_021E,
         MPop70_74 = B01001_022E,
         MPop75_79 = B01001_023E,
         MPop80_84 = B01001_024E,
         MPop85_plus = B01001_025E,
         FPop25_29 = B01001_035E,
         FPop30_34 = B01001_036E,
         FPop35_39 = B01001_037E,
         FPop40_44 = B01001_038E,
         FPop45_49 = B01001_039E,
         FPop50_54 = B01001_040E,
         FPop55_59 = B01001_041E,
         FPop60_61 = B01001_042E,
         FPop62_64 = B01001_043E,
         FPop65_66 = B01001_044E,
         FPop67_69 = B01001_045E,
         FPop70_74 = B01001_046E,
         FPop75_79 = B01001_047E,
         FPop80_84 = B01001_048E,
         FPop85_plus = B01001_049E,
         FBach25_34 = B15001_058E,
         FBach35_44 = B15001_066E,
         FBach45_64 = B15001_074E,
         FBach65plus = B15001_082E,
         MBach25_34 = B15001_017E,
         MBach35_44 = B15001_025E,
         MBach45_64 = B15001_033E,
         MBach65plus = B15001_041E,
         MedHHInc = B19013_001E, 
         MedRent = B25058_001E,
         TotalPoverty = B06012_002E) %>%
  dplyr::select(-NAME, -starts_with("B")) %>%
  mutate(pctWhite = ifelse(TotalPop > 0, Whites / TotalPop, 0),
         TotalMPop25plus = MPop25_29 + MPop30_34 + MPop35_39 + MPop40_44 + MPop45_49 + MPop50_54 + MPop55_59 + MPop60_61 + MPop62_64 + MPop65_66 + MPop67_69 + MPop70_74 + MPop75_79 + MPop80_84 + MPop85_plus,
         TotalFPop25plus = FPop25_29 + FPop30_34 + FPop35_39 + FPop40_44 + FPop45_49 + FPop50_54 + FPop55_59 + FPop60_61 + FPop62_64 + FPop65_66 + FPop67_69 + FPop70_74 + FPop75_79 + FPop80_84 + FPop85_plus,
         TotalPop25plus = TotalMPop25plus + TotalFPop25plus,
         FemaleBachelors = FBach25_34 + FBach35_44 + FBach45_64 + FBach65plus,
         MaleBachelors = MBach25_34 + MBach35_44 + MBach45_64 + MBach65plus,
         pctBachelors = ifelse(TotalPop > 0, ((FemaleBachelors + MaleBachelors) / TotalPop25plus), 0),
         pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
         year = "2022") %>%
  dplyr::select(GEOID,TotalPop,MedHHInc,MedRent,pctWhite,pctBachelors,pctPoverty,year)
```

*What is the value of combining data in this structural form??*

```{r}
allTracts <- rbind(tracts12,tracts22)

```

### Wrangling Transit Open Data

```{r}

DC_stop <- st_read("https://maps2.dcgis.dc.gov/dcgis/rest/services/DCGIS_DATA/Transportation_Rail_Bus_WebMercator/MapServer/52/query?outFields=*&where=1%3D1&f=geojson")

DC_line <- st_read("https://maps2.dcgis.dc.gov/dcgis/rest/services/DCGIS_DATA/Transportation_Rail_Bus_WebMercator/MapServer/106/query?outFields=*&where=1%3D1&f=geojson")

#Broad_St <- st_read("https://opendata.arcgis.com/datasets/2e9037fd5bef406488ffe5bb67d21312_0.geojson")
DC_stop <- st_transform(DC_stop, crs = 'ESRI:102728')

```

```{r}
library(mapview)
mapview(DC_stop)
```

Let's visualize it

```{r}
#septaStops <- 
  #rbind(
     #el %>% 
      #mutate(Line = "El") %>%
      #dplyr::select(Station, Line),
     #Broad_St %>%
      #mutate(Line ="Broad_St") %>%
      #dplyr::select(Station, Line)) %>%
  #st_transform(st_crs(tracts16))  

#this code use to combine
```

```{r}
ggplot() + 
  geom_sf(data=st_union(tracts12)) +
  geom_sf(data=DC_stop, 
          #aes(colour = LINE), 
          show.legend = "point", size= 2) +
  #scale_colour_manual(values = c("blue","red","orange")) +
  labs(title="DC_stop", 
       subtitle="Washington, DC", 
       caption="Figure 2.5") +
  mapTheme()
```

```{r}
ggplot() + 
  geom_sf(data=st_union(tracts12)) +
  geom_sf(data=DC_line, 
          aes(colour = NAME), 
          show.legend = "point", size= 2) +
  scale_colour_manual(values = c("blue","green","orange","red","black","yellow")) +
  labs(title="DC_line", 
       subtitle="Washington, DC", 
       caption="Figure 2.5") +
  mapTheme()
```

### Relating SEPTA Stops and Tracts

```{r}
stopBuffer <- st_buffer(DC_stop, 2640)# 2640 is understood as feet, half mile (2,640 ft.) of stations

stopUnion <- st_union(st_buffer(DC_stop, 2640))

DCBuffers <- 
  rbind(
     stopBuffer %>%
      mutate(Legend = "Buffer") %>%
      dplyr::select(Legend),
     stopUnion %>%
      st_sf() %>%
      mutate(Legend = "Unioned Buffer"))
```

Let's examine both buffers by making a "[small
multiples](https://www.juiceanalytics.com/writing/better-know-visualization-small-multiples)"
plot. This type of plot shows multiple views of the same data or
geographic area. We accomplish this by using the `facet_wrap()`
{ggplot2} function on the *Legend* column.

```{r}
ggplot() +
  geom_sf(data=DCBuffers) +
  geom_sf(data=DC_stop, show.legend = "point") +
  facet_wrap(~Legend) + 
  labs(caption = "Figure 2.6", title = "half mile buffers") +
  mapTheme()
```

In SE area, there have largest transit stop density, which has many
Federal building and landmarkin this area. More transit stop in this
area that convinience for people traveling.

### Using the `sf` Package for Spatial operations

```{r}
buffer <- filter(DCBuffers, Legend=="Unioned Buffer")

```


### Spatial Intersection of tracts and buffers

#### Spatial Clipping with `st_intersection()` on polygons

```{r}
clip <- 
  st_intersection(buffer, tracts12) %>%
  dplyr::select(TotalPop, MedRent) %>%
  mutate(Selection_Type = "Clip")
```

#### Spatial intersection with `st_intersects()` on polygons

```{r spatialSelection}

# Do a spatial selection to see which tracts touch the buffer

# approach #1: sub-setting a spatial object with a spatial object using the '[' brackets.
selection1 <- 
  tracts12[buffer,] %>%
  dplyr::select(TotalPop, MedRent) %>%
  mutate(Selection_Type = "Spatial Selection")

# approach #2: using `st_intersects` as a verbose way to do approach #1
selection2 <- tracts12[st_intersects(tracts12, buffer) %>% lengths > 0, ] %>%
  dplyr::select(TotalPop,MedRent) %>%
  mutate(Selection_Type = "Spatial Selection")

# approach #3: use `st_join` to do a spatial join and remove the non-intersecting polygons
selection3 <- tracts12 %>% 
  st_join(buffer, join = st_intersects) %>% 
  filter(!is.na(Legend)) %>% 
  dplyr::select(TotalPop, MedRent) %>%
  mutate(Selection_Type = "Spatial Intersects")
```


#### Spatial intersection with with `st_centroid()` on polygon centroids

```{r}
selectCentroids <-
  st_centroid(tracts12)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts12, GEOID), by = "GEOID") %>%
  st_sf() %>%
  dplyr::select(TotalPop, MedRent) %>%
  mutate(Selection_Type = "Select by Centroids")
  
```

summarise(Pop = mean(TotalPop, na.rm = TRUE),
            Rent = mean(MedRent, na.rm = TRUE))%>%

Let's plot the results of each method to assess.

```{r}
view(selectCentroids)
```

```{r}
intersections <- rbind(clip, selection1, selectCentroids)

ggplot() +
  geom_sf(data=intersections, aes(fill = TotalPop)) +
  geom_sf(data=DC_stop, show.legend = "point") +
  scale_fill_viridis_c() +
  facet_wrap(~Selection_Type) + 
  mapTheme()

```

```{r}
selectCentroids_12 <-
  st_centroid(tracts12)[buffer,] %>%
  st_drop_geometry() %>%
  left_join(., dplyr::select(tracts12, GEOID), by = "GEOID") %>%
  st_sf() %>%
  group_by() %>%
   summarise(Pop = mean(TotalPop, na.rm = TRUE),
            Rent = mean(MedRent, na.rm = TRUE)) %>%
  dplyr::select(Pop, Rent) %>%
  mutate(lon = st_coordinates(geometry)[,1],
         lat = st_coordinates(geometry)[,2])

#resulting to station
#DC_centroid <- cbind(DCBuffers, selectCentroids_12) 
```
```{r}
 # Spatial join: Find the intersection between buffers and tracts
buffer_population <- st_intersection(stopBuffer, tracts12)

# Calculate the population within each buffer by summing the population in intersecting tracts
buffer_population_summary <- st_sf(buffer_population ) %>%
  st_centroid() %>%
    mutate(lon = st_coordinates(geometry)[,1],
         lat = st_coordinates(geometry)[,2])

```

```{r}
ggplot() +
  geom_sf(data=allTracts,show.legend = "point") +
  #geom_sf(data= buffer_population_summary, aes(fill = total_population)) +
  geom_point(data = buffer_population_summary,
             aes(x= lon,
                 y= lat,
                 size = TotalPop),
             color = "purple", alpha = 0.5)  +
  scale_fill_viridis_c() +
  #facet_wrap(~Selection_Type) + 
  labs(title = "population and half mile buffers of each Metro station") +
  mapTheme()
    
```



```{r}
ggplot() +
  geom_sf(data=allTracts,show.legend = "point") +
  geom_sf(data= buffer_population_summary,aes(fill = Rent)) +
    geom_sf(data = DC_stop,
          fill = alpha("red", 0.7),
          col = "grey20")+
    scale_fill_manual(values = palette5,labels = qBr(buffer_population_summary, "Rent"),
                    name = "Rent\n(Quintile Breaks)") +
  scale_fill_viridis_c() +
  #facet_wrap(~Selection_Type) + 
  labs(title = "Rent and half mile buffers of each Metro station") +
  mapTheme()
```


```{r}
ggplot(myData)+
  geom_point(data = st_union(tracts12))+
  geom_sf(aes(fill = q5(TotalPop))) +
  geom_sf(data = DC_stop,
          fill = alpha("red", 0.7),
          col = "grey20")+
  scale_fill_manual(values = palette5,
                    labels = qBr(myData, "TotalPop"),
                    name = "Popluation\n(Quintile Breaks)") +
  labs(title = "Total Population", subtitle = "DC; 2012") +
  facet_wrap(~Selection_Type)+
  mapTheme() + 
  theme(plot.title = element_text(size=22))
```


```{r}
ggplot() +
  geom_sf(data=allTracts,show.legend = "polygon") +
  geom_sf(data= intersections, show.legend = "polygon", aes(fill = MedRent)) +
  scale_fill_viridis_c() +
  facet_wrap(~Selection_Type) + 
  labs(title = "Rent and half mile buffers of each Metro station") +
  mapTheme()
```

```{r}
ggplot(myData)+
  geom_sf(data = st_union(tracts12))+
  geom_sf(aes(fill = q5(MedRent))) +
  scale_fill_manual(values = palette5,
                    labels = qBr(myData, "MedRent"),
                    name = "Popluation\n(Quintile Breaks)") +
  labs(title = "Mean Rent", subtitle = "DC; 2012") +
  facet_wrap(~Selection_Type)+
  mapTheme() + 
  theme(plot.title = element_text(size=22))
```

### Indicator Maps

```{r}
allTracts.group <- 
  rbind(
    st_centroid(allTracts)[buffer,] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(allTracts)[buffer, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
  mutate(MedRent.inf = ifelse(year == "2012", MedRent * 1.14, MedRent)) 

```

```{r}
ggplot() +
  #geom_sf(data=allTracts,show.legend = "polygon") +
  geom_sf(data= allTracts.group,  aes(color = TOD)) +
  #scale_color_discrete(name = "TOD") +
  scale_fill_viridis_c() +
  facet_wrap(~year) + 
  labs(title = "time and space (TOD vs. non-TOD)") +
  mapTheme()

```

```{r}
  ggplot(allTracts.group)+
    geom_sf(data = st_union(tracts12))+
    geom_sf(aes(fill = TOD)) +
    labs(title = "Time/Space Groups") +
    facet_wrap(~year)+
    mapTheme() + 
    theme(plot.title = element_text(size=22))
```

```{r}
  ggplot(allTracts.group)+
    geom_sf(data = st_union(tracts12))+
    geom_sf(aes(fill = q5(MedRent.inf))) +
    geom_sf(data = buffer, fill = "transparent", color = "red")+
    scale_fill_manual(values = palette5,
                      labels = qBr(allTracts.group, "MedRent.inf"),
                      name = "Rent\n(Quintile Breaks)") +
    labs(title = "Median Rent 2012-2022", subtitle = "Real Dollars") +
    facet_wrap(~year)+
    mapTheme() + 
    theme(plot.title = element_text(size=22))
```

*Can you try to create the small multiple indicator maps seen in the
text? The solutions are contained in "map_exercise.R"*

## Appendices - More Code For Completing Assignment 2

### TOD Indicator Tables

Your homework will require you to make indicator tables comparing TOD
and non TOD areas over time. These code blocks create the tables you see
in the book and in lecture.

```{r}
allTracts.Summary <- 
  st_drop_geometry(allTracts.group) %>%
  group_by(year, TOD) %>%
  summarize(Rent = mean(MedRent, na.rm = T),
            Population = mean(TotalPop, na.rm = T),
            Percent_White = mean(pctWhite, na.rm = T),
            Percent_Bach = mean(pctBachelors, na.rm = T),
            Percent_Poverty = mean(pctPoverty, na.rm = T))

kable(allTracts.Summary) %>%
  kable_styling() %>%
  footnote(general_title = "\n",
           general = "Table 2.2")
```

Let's make some comparisons and speculate about the willingness to pay
and demographics in these areas 2016-2020 (see the 2000 data in the text
too)

Notice how we pipe the kable() command here

```{r}
allTracts.Summary %>%
  unite(year.TOD, year, TOD, sep = ": ", remove = T) %>%
  gather(Variable, Value, -year.TOD) %>%
  mutate(Value = round(Value, 2)) %>%
  spread(year.TOD, Value) %>%
  kable() %>%
  kable_styling() %>%
  footnote(general_title = "\n",
           general = "Table 2.3")
```

### TOD Indicator Plots

```{r}
allTracts.Summary %>%
  gather(Variable, Value, -year, -TOD) %>%
  ggplot(aes(year, Value, fill = TOD)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Variable, scales = "free", ncol=5) +
  scale_fill_manual(values = c("#bae4bc", "#0868ac")) +
  labs(title = "Indicator differences across time and space" )+
  plotTheme() + 
  theme(legend.position="bottom")
```




```{r}
allTracts.DCBuffers <-
  st_join(st_centroid(allTracts), DCBuffers) %>%
  st_drop_geometry() %>%
  left_join(allTracts) %>%
  #mutate(Submarket = replace_na(Submarket, "Non-TOD")) %>%
  st_sf() 
```
```{r}
allTracts.group <- 
  rbind(
    st_centroid(allTracts)[buffer,] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "TOD"),
    st_centroid(allTracts)[buffer, op = st_disjoint] %>%
      st_drop_geometry() %>%
      left_join(allTracts) %>%
      st_sf() %>%
      mutate(TOD = "Non-TOD")) %>%
  mutate(MedRent.inf = ifelse(year == "2012", MedRent * 1.42, MedRent)) 
```

#### Example of using `multipleRingBuffer()` function

Create a geom_line plot that shows mean rent as a function of distance
to subway stations (Figure 1.17). To do this you will need to use the
multipleRingBuffer function found in the functions.R script. Check out
the MRB_Explained markdown on Canvas for more info on how this works!

```{r}
DC_stop_MRB <- multipleRingBuffer(st_union(DC_stop),
                                maxDistance = 47520,
                                interval =  2640)
```

```{r}
allTracts.rings <-
  st_join(st_centroid(dplyr::select(allTracts, GEOID, year)), 
          multipleRingBuffer(st_union(DC_stop), 47520, 2640)) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts, GEOID, MedRent, year), 
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance = distance / 5280) #convert to miles
```

```{r}
ggplot() +
    geom_sf(data=DC_stop_MRB) +
    geom_sf(data=DC_stop, size=1) +
    geom_sf(data=st_union(tracts22), fill=NA, size=1.2) +
    labs(title="Half mile buffers") +
    mapTheme()

```


```{r}
ggplot() +
    geom_sf(data=DC_stop_MRB) +
    geom_sf(data=DC_stop, size=1) +
   geom_sf(data=allTracts, aes(fill = TotalPop)) +
    labs(title="Half mile buffers") +
  
    mapTheme()

```

```{r}
allTracts.rings <- allTracts %>% 
  select(GEOID, year) %>% 
  st_centroid() %>% 
  st_join(DC_stop_MRB, join = st_intersects) %>%
  st_drop_geometry() %>%
  left_join(dplyr::select(allTracts, GEOID, MedRent, year), 
            by=c("GEOID"="GEOID", "year"="year")) %>%
  st_sf() %>%
  mutate(distance = distance / 5280) #convert to miles

```

```{r}
allTracts.rings.summary <- st_drop_geometry(allTracts.rings) %>%
    group_by(distance, year) %>%
    summarize(Mean_Rent = mean(MedRent, na.rm=T))
```

```{r}
ggplot(allTracts.rings.summary,
       aes(distance, Mean_Rent, colour=year)) +
      geom_point(size=3) + 
  labs(title="Rent as function of distance to Metro station") +
  geom_line(size=2)
```

Comparing two different year, the rent have increasing around 1k during
last ten years. However, there has been no significant change in the
relationship between rent and distance to Metro stations.
